---
title: "ICC: cross-run (split-half), ub55, cross-task projection onto hilo axis"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true

---


# intro


```{r setup, include = FALSE}

library(here)
library(here)
library(tidyr)
library(dplyr)
library(data.table)
library(abind)
library(doParallel)
library(foreach)
library(mikeutils)
library(progress)
library(ggplot2)
library(grid)
library(gridExtra)
library(cowplot)
library(viridis)

theme_set(theme_half_open())

source(here("code", "_constants.R"))
source(here("code", "_atlases.R"))
source(here("code", "_funs.R"))

fnames <- list.files(here("out", "icc"), full.names = TRUE)
projs <- setNames(lapply(fnames, readRDS), list.files(here("out", "icc")))
projs <- lapply(projs, reshape::melt) %>% lapply(data.table)
d <- rbindlist(projs, fill = TRUE, idcol = "file")

d <- d %>% filter(!is.na(value))


```



# within-task projections

```{r}

d_wthn <- d %>% 
  
  filter(grepl("wthntsk_parc", file)) %>% 
  select(-run) %>%
  mutate(
    file = gsub("hiloaxis-projections-wthntsk_parc-schaefer07_prew-concat-runs-|.RDS", "", file),
    standardized = grepl("axes", file)
  )

d_wthn <- bind_rows(
  
  d_wthn %>%
    filter(method == "prewhitened") %>%
    mutate(prewhitened = substr(file, 1, 4)),
  
  d_wthn %>%
    filter(method == "vanilla", grepl("multi", file)) %>%
    mutate(prewhitened = "none")
  
  ) %>% select(-file, -method)


d_wthn %>%
  
  filter(standardized, prewhitened %in% c("none", "univ")) %>%
  
  group_by(standardized, prewhitened, task, roi) %>%
  mutate(value = value / sd(value)) %>%

  ggplot(aes(prewhitened, value, fill = prewhitened)) +
  geom_hline(yintercept = 0) +
  
  stat_summary(fun = "mean", geom = "col", position = position_dodge(width = 1)) +
  stat_summary(
    fun.data = "mean_cl_boot", geom = "errorbar", width = 0, size = 1.5,
    position = position_dodge(width = 1)
    ) +
  
  facet_grid(vars(task), vars(roi)) +
  scale_fill_brewer(type = "qual", palette = 6)

## prewhitening seems to reduce variability across-the-board (all tasks)
## massive increases in CTS and Stern.... (and increase raw mean effect, too...)
## axcpt and stroop, much less strong of impact, but seems to generally increase std ES
## really not sure what's going on with CTS and Sternberg.
## should investigate noise model GLM and noise model code more...


d_sum <-
  d_wthn %>%
  
  filter(standardized) %>%

  pivot_wider(
    id_cols = c("standardized", "subj", "roi", "prewhitened"), 
    names_from = "task", values_from = "value"
    ) %>%
  group_by(standardized, prewhitened, roi) %>%
  
  summarize(
    r_axcpt_cuedts = cor(Axcpt, Cuedts),
    r_axcpt_stern = cor(Axcpt, Stern),
    r_axcpt_stroop = cor(Axcpt, Stroop),
    r_cuedts_stern = cor(Cuedts, Stern),
    r_cuedts_stroop = cor(Cuedts, Stroop),
    r_stern_stroop = cor(Stern, Stroop)
    ) %>%
    
    View


```



# between-task

## cross-run correlation (reliability)

```{r}

r <-
  d %>%
  
  filter(!grepl("wthntsk", file)) %>%
  pivot_wider(id_cols = c("subj", "roi", "task", "file"), names_from = "run", values_from = "value") %>%
  
  group_by(roi, task, file) %>%
  summarize(r = cor(run1, run2))


r %>%
  
  filter(grepl("means|stand-axes_cvweight-unif", file)) %>%
  
  ggplot(aes(file, r, fill = file)) +
  geom_hline(yintercept = 0) +
  
  geom_col(position = position_dodge(width = 1)) +

  facet_grid(vars(task), cols = vars(roi)) +
  scale_fill_brewer(type = "qual", palette = 6) +
  theme(axis.text.x = element_blank()) +
  
  labs(title = "cross-run correlation")


## univariate noise norm and non-prewhitened seem to have best reliability
## max is 0.5 (axcpt, control network)
## univariate mean quite poor reliability -- max ~ 0.2


```


# group-level effects

```{r}


d %>%
  
  filter(grepl("means|none_stand-axes_cvweight-unif|univ_stand-axes_cvweight-unif", file)) %>%
  
  group_by(file, task, roi, method, subj) %>%
  summarize(value = mean(value)) %>%  ## average across runs
  mutate(
    value = value / sd(value),
    file = gsub("hiloaxis-projections_parc-schaefer07_prew-|.RDS", "", file)
    ) %>%
  
  ggplot(aes(file, value, fill = file)) +
  geom_hline(yintercept = 0) +
  
  stat_summary(fun = "mean", geom = "col", position = position_dodge(width = 1)) +
  stat_summary(
    fun.data = "mean_cl_boot", geom = "errorbar", width = 0, size = 1.5,
    position = position_dodge(width = 1)
    ) +
  
  facet_grid(vars(task), vars(roi)) +
  scale_fill_brewer(type = "qual", palette = 6) +
  theme(axis.text.x = element_blank())


## univ and no-prewhitening have comparable impact on group-level effect size for cross-task axis.


# d %>%
#   
#   filter(grepl("means|none_stand-axes_cvweight-unif", file)) %>%  
#   
#   group_by(file, task, roi, method, subj) %>%
#   summarize(value = mean(value)) %>%
#   mutate(
#     # value = value / sd(value),
#     file = gsub("hiloaxis-projections_parc-schaefer07_prew-|.RDS", "", file)
#     ) %>%
#   
#   pivot_wider(id_cols = c("subj", "file", "task"), names_from = "roi", values_from = "value") %>%
#   group_by(file, task) %>%
#   
#   filter(
#     grepl("univ_stand-axes_cvweight-unif", file), task == "Stroop"
#   ) %>%
#   ungroup %>%
#   select(Cont:Vis)  %>%
#   cor %>%
#   matplot + 
#   theme(legend.position = "top", axis.text = element_text()) + 
#   geom_text(aes(label = round(value, 1)), color = "grey20") +
#   scale_fill_continuous(limits = c(0, 1))
# 
# 



```



# quick look

```{r}

d %>%
  
  ggplot(aes(value)) +
  geom_histogram()


d %>%
  
  ggplot(aes(d)) +
  geom_histogram()


d %>%
  
  ggplot(aes(d)) +
  geom_histogram() +
  facet_grid(vars(parcel), vars(task))
  

```


# group-level effects

Are projections generally positive across subjs?
That is, was "hi" vs "lo" successfully decoded?


```{r}

d %>%
  
  ggplot(aes(run, d, fill = run)) +
  geom_hline(yintercept = 0) +
  
  stat_summary(fun = "mean", geom = "col") +
  stat_summary(fun.data = "mean_cl_boot", geom = "errorbar", width = 0, size = 1.5) +
  
  facet_grid(vars(task), vars(parcel)) +
  scale_fill_brewer(type = "qual", palette = 6) +
  theme(legend.position = "none") +
  
  labs(
    y = "mean cross-task projection on hi-lo axis\n(cohen's D)",
    x = "scanning run",
    caption = "error bars: 95% CI (bootstrapped)"
  )


sum_stats <- d %>%
  
  group_by(task, parcel) %>%
  summarize(
    statistic = t.test(d)$statistic,
    p = t.test(d)$p.value
  )

sum_stats %>% filter(p < 0.001)

```



# individual effects

Is an individual's projection generally stable across scanning runs of the same task?
That is, what is the split-half reliability of "hi" vs "lo" multivariate contrast?


```{r}

d_w <- d %>% 
  pivot_wider(id_cols = c("task", "parcel", "subj"), names_from = "run", values_from = "d")
  
d_w %>%
  
  ggplot(aes(run1, run2)) +
  geom_abline() +
  
  geom_point(shape = 21, color = "white", fill = "black", size = 3) +
  
  facet_grid(vars(task), vars(parcel)) +
  
  labs(
    x = "run 1 cross-task projection on hi-lo axis\n(cohen's D)",
    y = "run 2 cross-task projection on hi-lo axis\n(cohen's D)",
    caption = "error bars: 95% CI (bootstrapped)"
  )


sum_stats_indiv <- d_w %>%
  
  group_by(task, parcel) %>%
  summarize(
    r = cor(run1, run2),
    rho = cor(run1, run2, method = "spearman")
  )

sum_stats_indiv %>% arrange(-r)

```


