---
title: "Group analysis, runwise GLMs"
author: "mike freund"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    highlight: zenburn
---


```{r setup, include = FALSE}

knitr::opts_chunk$set(
  cache = TRUE, echo = FALSE, warning = FALSE, message = FALSE,
  fig.align = 'center',
  fig.width = 11.5, fig.fullwidth = TRUE
  )

library(here)
library(magrittr)
library(gifti)
library(cifti)
library(abind)
library(mikeutils)
library(lme4)
library(ggplot2)
library(ggbeeswarm)
library(dplyr)
library(data.table)
library(tawny)  ## shrinkage.intensity()
library(expm)  ## sqrtm()
library(grid)
library(gridExtra)
library(cowplot)

theme_set(theme_classic(base_size = 8))

r <- readRDS(here("out", "runwise", "qc_group", "reliability_hilo_baseline_schaefer400.rds"))
d <- readRDS(here("out", "runwise", "qc_group", "estimates-win-run_hilo_baseline_schaefer400.rds"))

nodename <- Sys.info()["nodename"]
if (nodename == "ccplinux1") {
  dir.atlas <- "/data/nil-external/ccp/freund/atlases"
} else if (nodename == "CCP-FREUND") {
  dir.atlas <- "C:/local/atlases"
}

parcellation <- read_atlas("schaefer400")
hcp <- list(
  L  = readGIfTI(
    file.path(dir.atlas, "surf", "HCP_S1200_GroupAvg_v1", "S1200.L.inflated_MSMAll.32k_fs_LR.surf.gii")
    ),
  R = readGIfTI(
    file.path(dir.atlas, "surf", "HCP_S1200_GroupAvg_v1", "S1200.R.inflated_MSMAll.32k_fs_LR.surf.gii")
    )
)
over <- list(
  L = parcellation$atlas[1:(nrow(parcellation$atlas) / 2)], 
  R = parcellation$atlas[(nrow(parcellation$atlas) / 2):nrow(parcellation$atlas)]
  )

```


```{r check_data, include = FALSE}

dimnames(r)
dimnames(d)
u <- d[, , , "mean", ]  ## u for univariate

subj.has.na.r <- apply(r, "subj", function(.) any(is.na(c(.))))
subj.has.na.u <- apply(u, "subj", function(.) any(is.na(c(.))))
identical(subj.has.na.u, subj.has.na.r)

subj.all.data <- names(subj.has.na.r)[!subj.has.na.r]

r <- r[subj.all.data, , , ]
u <- u[subj.all.data, , , ]

## create dataframes

r.df <- reshape2::melt(r)
r.df <- cbind(r.df, reshape2::colsplit(r.df$roi, "_", c("hemi", "community", "parcel")))
r.df <- purrr::map_if(r.df, is.factor, as.character) %>% as.data.frame

r.df %<>%
  group_by(roi) %>%
  mutate(roi.num = match(roi, parcellation$key))
str(r.df)


u.df <- reshape2::melt(u)
u.df <- cbind(u.df, reshape2::colsplit(u.df$roi, "_", c("hemi", "community", "parcel")))
u.df <- purrr::map_if(u.df, is.factor, as.character) %>% as.data.frame

u.df %<>%
  group_by(roi) %>%
  mutate(roi.num = match(roi, parcellation$key))

u.wide <- u.df %>% tidyr::spread(run, value)


```

## 'activation' analysis: group-mean univariate contrast between hi-lo

Parcels are highlighted with abs(t) > 2.

```{r activation, fig.height = 7}

u.df.byroi <- split(u.df, u.df$roi)

mods.activation <- lapply(u.df.byroi, function(.) lmer(value ~ task - 1 + (1 | subj), .))
coef.activation <- lapply(mods.activation, function(.) as.data.frame(coef(summary(.))))
coef.activation %<>% lapply(function(.) tibble::rownames_to_column(., var = "task"))
coef.activation <- as.data.frame(do.call(rbind, coef.activation))
coef.activation$roi <- rep(names(mods.activation), each = 4)
coef.activation$task <- gsub("task", "", coef.activation$task)

coef.activation %<>% cbind(reshape2::colsplit(.$roi, "_", c("hemi", "community", "parcel")))

coef.activation %<>% rename(b = "Estimate", t = "t value", se = "Std. Error")

coef.activation %>%
  ggplot(aes(parcel, b)) +
  facet_grid(vars(task), vars(community), scales = "free", space = "free") +
  geom_hline(yintercept = 0) +
  geom_point(aes(alpha = ifelse(abs(t) > 2, 1, 1/12))) +
  geom_errorbar(
    aes(
      ymin = b - se * 1.96, ymax = b + se * 1.96,
      alpha = ifelse(abs(t) > 2, 1, 1/12)
      )
    ) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, color = ifelse(abs(coef.activation[["t"]]) > 2, "black", "grey70")),
    legend.position = "none"
  )

```

```{r activation_brains, fig.height = 4, dev = "png", dpi = 72}

build.overlay <- function(df, column) {
  
  df <- df[match(parcellation$key, df$roi), ]
  df <- df[match(parcellation$key, df$roi), ]
  
  for (i in seq_along(over)) over[[i]][over[[i]] == 0] <- NA
  over <- lapply(over, function(.) df[., column])
  for (i in seq_along(over)) over[[i]][is.na(over[[i]])] <- 0
  
  over
  
}

tasks <- as.character(unique(r.df$task))
subjs <- subj.all.data

for (task.i in seq_along(tasks)) {
  
  plot_surface(
    build.overlay(filter(coef.activation, task == tasks[task.i]), "t"),
    hcp
  )
  mtext(title(tasks[task.i]), side = 3, line = 0, outer = TRUE)
  
}

```

## reliability analysis

#### Stats

For a given parcel and task...

__Univariate reliability__:

  * the between-subject correlation of mean hi--lo estimates across scanning runs: $r^{\text{(univ)}}$
  * bootstrapped 95% confidence interval around $r^{\text{(univ)}}$ (bias-corrected and accelerated)

__Multivariate reliability__:

  * a vector of within-subject correlations of hi--lo patterns across scanning runs: $r^{\text{(multiv)}}_{s}$
  * take mean correlation across subjects: $\bar r^{\text{(multiv)}} = \sum_{s = 1}^{N_{subjs}} r^{\text{(multiv)}}_s/N_{\text{subjs}}$ (using z-transformed measures)
  * bootstrapped 95% confidence interval (bias-corrected and accelerated)

#### Figure

  * **BLUE**: parcels with $r^{\text{(univ)}} > 0$ (lower bound of 95% CI)
  * **RED**: parcels with $\bar r^{\text{(multiv)}} > 0$ (lower bound of 95% CI)
  * **PURPLE**: parcels with $r^{\text{(univ)}} > 0$ and $\bar r^{\text{(multiv)}} > 0$

```{r, reliability, fig.height = 7}

## bootstrap confidence intervals for comparison plot

if (file.exists(here("luniv.RDS")) && file.exists(here("lmult.RDS"))) {

  l.univ <- readRDS(here("luniv.RDS"))
  l.mult <- readRDS(here("lmult.RDS"))

} else {

  # splithalf <- function(x, ii) cor(x[ii, "run1"], x[ii, "run2"])
  splithalf <- function(x, ii) cor(x[ii, ])[1, 2]
  mean.rzr <- function(x, ii) tanh(mean(atanh(x[ii])))

  bootci <- function(x, f, R) {
    b <- boot::boot(x, f, R)
    boot::boot.ci(b, type = "bca")$bca[4:5]
  }


  r.df <- as.data.table(r.df)[stat == "b"]
  u.wide <- as.data.table(u.wide)
  l.univ <- setNames(vector("list", length(parcellation$key) * length(tasks)), combo_paste(parcellation$key, tasks))
  l.mult <- l.univ

  for (roi.i in seq_along(parcellation$key)) {
    for (task.j in seq_along(tasks)) {

      name.roi.i <- parcellation$key[roi.i]
      name.task.j <- tasks[task.j]

      v <- r.df[roi == name.roi.i & task == name.task.j]$value
      m <- u.wide[roi == name.roi.i & task == name.task.j, c("run1", "run2")]

      name.ij <- paste0(name.roi.i, "_", name.task.j)
      l.univ[[name.ij]] <- bootci(v, splithalf, R = 1000)
      l.mult[[name.ij]] <- bootci(m, mean.rzr, R = 1000)

    }

    print(roi.i)

  }

  saveRDS(l.univ, here("luniv.RDS"))
  saveRDS(l.mult, here("lmult.RDS"))

}

## some ugly wrangling

# cis <- as.data.frame(cbind(do.call(rbind, l.univ), do.call(rbind, l.mult)))
# names(cis) <- c("l.univ", "h.univ", "l.mult", "h.mult")
# cis$task <- lapply(
#   tasks,
#   function(x) ifelse(grepl(x, names(l.univ)), x, "")
#   ) %>%
#   do.call(cbind, .) %>%
#   apply(1, paste0, collapse = "")
# cis$roi <- gsub(paste0("_", tasks, collapse = "|"), "", names(l.univ))


ci.univ <- as.data.frame(do.call(rbind, l.univ))
names(ci.univ) <- c("lb", "ub")
ci.univ$task <- lapply(
  tasks,
  function(x) ifelse(grepl(x, names(l.univ)), x, "")
  ) %>%
  do.call(cbind, .) %>%
  apply(1, paste0, collapse = "")
ci.univ$roi <- gsub(paste0("_", tasks, collapse = "|"), "", names(l.univ))

ci.mult <- as.data.frame(do.call(rbind, l.mult))
names(ci.mult) <- c("lb", "ub")
ci.mult$task <- ci.univ$task
ci.mult$roi <- ci.univ$roi

# u.group <- u.wide %>%
#   group_by(task, roi) %>%
#   summarize(r = cor(run1, run2)) %>%
#   full_join(ci.univ, by = c("task", "roi")) %>%
#   mutate(method = "univ") %>%
#   as.data.frame
#
# r.group <- r.df %>%
#   filter(stat == "b") %>% select(-stat) %>%
#   group_by(task, roi) %>%
#   summarize(
#     r = tanh(mean(atanh(value))),
#     p    = wilcox.test(value, alternative = "greater")$p.value
#     ) %>%
#   mutate(p.adj = p.adjust(p, method = "fdr")) %>%
#   full_join(ci.mult, by = c("task", "roi")) %>%
#   mutate(method = "mult")
#
# group <- as.data.frame(bind_rows(r.group, u.group))
# group <- cbind(group, reshape2::colsplit(group$roi, "_", c("hemi", "community", "parcel")))

u.group <- u.wide %>%
  group_by(task, roi) %>%
  summarize(r.univ = cor(run1, run2)) %>%
  full_join(ci.univ, by = c("task", "roi")) %>%
  as.data.frame

r.group <- r.df %>%
  filter(stat == "b") %>% select(-stat) %>%
  group_by(task, roi) %>%
  summarize(
    r.mult = tanh(mean(atanh(value))),
    p      = wilcox.test(value, alternative = "greater")$p.value
    ) %>%
  mutate(p.adj = p.adjust(p, method = "fdr")) %>%
  full_join(ci.mult, by = c("task", "roi"))

group <- as.data.frame(bind_rows(r.group, u.group))
group <- full_join(r.group, u.group, by = c("task", "roi"), suffix = c(".mult", ".univ"))

group <- cbind(as.data.frame(group), reshape2::colsplit(group$roi, "_", c("hemi", "community", "parcel")))

group %<>%
  mutate(
    sig = ifelse(
      lb.mult > 0 & lb.univ <= 0,
      "sigmult",
      # "#e34a33",
      ifelse(
        lb.univ > 0 & lb.mult <= 0,
        "siguniv",
        # "#3182bd",
        ifelse(
          lb.univ > 0 & lb.mult > 0,
          "sig",
          # "#8856a7",
          "ns"
        )
      )
    )
  )

group %>%
  ggplot(aes(r.univ, r.mult)) +
  facet_grid(vars(task), vars(community)) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_abline() +
  geom_errorbarh(aes(xmin = lb.univ, xmax = ub.univ), alpha = 1/12) +
  geom_errorbar(aes(ymin = lb.mult, ymax = ub.mult), alpha = 1/12) +
  geom_point(
    aes(fill = sig, alpha = ifelse(lb.mult < 0 & lb.univ < 0, 1/6, 1)),
    size = 2,
    shape = 21,
    color = "grey50"
    ) +
  scale_fill_manual(values = c(sigmult = "#e34a33", siguniv = "#3182bd", sig = "#8856a7", ns = "black")) +
  labs(
    title = "split-half reliability: univariate versus multivariate",
    x = "univariate (btw-subj)",
    y = "multivariate (win-subj)"
    ) +
  theme(
    legend.position = "none"
  ) +
  coord_cartesian(xlim = c(-0.6, 0.6), ylim = c(-0.6, 0.6))

# max(abs(c(range(group$r.univ), range(group$r.mult))))

```

#### univariate

Parcels are highlighted with lower-bound CIs > 0.

```{r, reliability_univ, fig.height = 7}

group %>%
  ggplot(aes(parcel, r.univ)) +
  facet_grid(vars(task), vars(community), scales = "free", space = "free") +
  geom_hline(yintercept = 0) +
  geom_point(aes(alpha = ifelse(lb.univ > 0, 1, 1/12))) +
  geom_errorbar(
    aes(
      ymin = lb.univ, ymax = ub.univ,
      alpha = ifelse(lb.univ > 0, 1, 1/12)
      )
    ) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, color = ifelse(group$lb.univ > 0, "black", "grey70")),
    legend.position = "none"
  )

```

#### multivariate

Parcels are highlighted with lower-bound CIs > 0.

```{r, reliability_mult, fig.height = 7}

group %>%
  ggplot(aes(parcel, r.mult)) +
  facet_grid(vars(task), vars(community), scales = "free", space = "free") +
  geom_hline(yintercept = 0) +
  geom_point(aes(alpha = ifelse(lb.mult > 0, 1, 1/12))) +
  geom_errorbar(
    aes(
      ymin = lb.mult, ymax = ub.mult,
      alpha = ifelse(lb.mult > 0, 1, 1/12)
      )
    ) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, color = ifelse(group$lb.mult > 0, "black", "grey70")),
    legend.position = "none"
  )

```


#### multivariate reliability: denoising methods


```{r, denoising, fig.height = 8}

activated <- coef.activation[abs(coef.activation$t) > 2, "roi"]
r.df$was.activated <- r.df$roi %in% activated

## scatterplots

n.resamples <- 1000

# r.df %>%
#   filter(stat %in% c("b", "t"), community %in% "Cont") %>%
#   tidyr::spread(stat, value) %>%
#   ggplot(aes(b, t, fill = community)) +
#   geom_abline() +
#   geom_hline(yintercept = 0) +
#   geom_vline(xintercept = 0) +
#   geom_point(shape = 21, color = "white") +
#   scale_fill_brewer(type = "qual", palette = 6) +
#   coord_cartesian(xlim = c(-1, 1), ylim = c(-1, 1))
# r.df %>%
#   filter(stat %in% c("b", "prewhitened"), community %in% "Cont") %>%
#   tidyr::spread(stat, value) %>%
#   ggplot(aes(b, prewhitened, fill = community)) +
#   geom_abline() +
#   geom_hline(yintercept = 0) +
#   geom_vline(xintercept = 0) +
#   geom_point(shape = 21, color = "white") +
#   scale_fill_brewer(type = "qual", palette = 6) +
#   coord_cartesian(xlim = c(-1, 1), ylim = c(-1, 1))


## "bar plots"

dodge.width <- 0.75

tasks <- as.character(unique(r.df$task))
communities <- unique(r.df$community)
p.parcels <- setNames(vector("list", length(communities)), communities)
p.tasks <- setNames(vector("list", length(tasks)), tasks)

for (task.i in seq_along(tasks)) {
  # task.i = 1

  for (community.i in seq_along(communities)) {
    # community.i = 5

    p <- r.df %>%
      filter(community == communities[community.i], task == tasks[task.i]) %>%
      ggplot(aes(parcel, value, color = stat, fill = stat)) +
      geom_hline(yintercept = 0) +
      # geom_quasirandom(
      #   dodge.width = dodge.width, shape = 21, size = 0.5, alpha = 1/12
      #   ) +
      stat_summary(
        aes(alpha = ifelse(was.activated, 1, 1/12)),
        fun.data = "mean_cl_boot", size = 1.25,
        geom = "errorbar",
        width = 0,
        fun.args = list(B = n.resamples),
        position = position_dodge(width = dodge.width)
      ) +
      scale_fill_brewer(type = "qual", palette = 2) +
      scale_color_brewer(type = "qual", palette = 2) +
      coord_flip() +
      theme(
        axis.line.y = element_blank(),
        legend.position = "none",
        axis.text.y = element_text(color = ifelse(r.df$was.activated, "black", "grey70"))
      ) +
      labs(title = communities[community.i], y = "correlation", x = "parcel")


    if (community.i == 1) {

      parcels <- r.df %>% filter(community == communities[community.i]) %>% .$parcel
      x.max <- length(unique(parcels))
      p <- p +
        annotate(
          "text", x = x.max, y = -Inf, label = 'bold("beta")', vjust = 0.5, hjust = 0, color = "#1b9e77",
          parse = TRUE
          ) +
        annotate(
          "text", x = x.max - 1, y = -Inf, label = 'bold("t stat")', vjust = 0.5, hjust = 0, color = "#d95f02",
          parse = TRUE
          ) +
        annotate(
          "text", x = x.max - 2, y = -Inf, label = 'bold("pre-whitened")', vjust = 0.5, hjust = 0, color = "#7570b3",
          parse = TRUE
          )

    } else {

      p <- p + theme(axis.title = element_blank())

    }

      p.parcels[[community.i]] <- p

  }

  p.tasks[[task.i]] <- plot_grid(
    p.parcels[[1]],
    p.parcels[[2]],
    p.parcels[[3]],
    p.parcels[[4]],
    p.parcels[[5]],
    p.parcels[[6]],
    p.parcels[[7]],
    align = "h",
    axis = "tb",
    nrow = 1
  )

}

p.tasks

```



```{r denoising_brains, fig.height = 4, dev = "png", dpi = 72}

r.group.all <- r.df %>%
  group_by(stat, task, roi, was.activated) %>%
  summarize(
    r.mult = tanh(mean(atanh(value))),
    p      = wilcox.test(value, alternative = "greater")$p.value
    ) %>%
  mutate(p.adj = p.adjust(p, method = "fdr"))

pairwise <- r.df %>%
  group_by(stat, task, roi, was.activated) %>%
  tidyr::spread(stat, value) %>%
  summarize(
    white.vs.b   = tanh(mean(atanh(prewhitened) - atanh(b))),
    t.vs.b       = tanh(mean(atanh(t) - atanh(b))),
    p.white.vs.b = wilcox.test(prewhitened, b, paired = TRUE)$p.value,
    p.t.vs.b     = wilcox.test(t, b, paired = TRUE)$p.value
  ) %>%
  mutate(
    p.adj.t.vs.b = p.adjust(t.vs.b, method = "fdr"),
    p.adj.white.vs.b = p.adjust(white.vs.b, method = "fdr")
    )

```

#### prewhitened versus "raw" betas

```{r denoising_brains_white_vs_raw, fig.height = 4, dev = "png", dpi = 72}

for (task.i in seq_along(tasks)) {
  
  df.task.i <- pairwise %>% filter(task == tasks[task.i]) %>% as.data.frame
  df.task.i$white.vs.b[df.task.i$p.white.vs.b > 0.05] <- 0
  
  plot_surface(
    build.overlay(df.task.i, "white.vs.b"),
    hcp
  )
  mtext(title(tasks[task.i]), side = 3, line = 0, outer = TRUE)
  
}

```

#### t-values versus "raw" betas

```{r denoising_brains_t_vs_raw, fig.height = 4, dev = "png", dpi = 72}

for (task.i in seq_along(tasks)) {
  
  df.task.i <- pairwise %>% filter(task == tasks[task.i]) %>% as.data.frame
  df.task.i$t.vs.b[df.task.i$p.t.vs.b > 0.05] <- 0
  
  plot_surface(
    build.overlay(df.task.i, "t.vs.b"),
    hcp
  )
  mtext(title(tasks[task.i]), side = 3, line = 0, outer = TRUE)
  
}

```

